---
title: "PA1_template"
author: "Mal"
date: "Thursday, January 15, 2015"
output: html_document
---

This is peer assesment 1 for Reproducible Research course on Coursera.
Here I will analyze some walking data and using knitr.  The end result is a clean and organized description of both the results I have found along with the code that I used to get it.  


First I loaded in my data.

```{r}
mydata <- read.csv("activity.csv", header = TRUE)
```

Then I had to transform and manipulate the data to analyze it.  First, I created a subset called completedata that had only the complete cases in it.  Then I split the data by date so I could use the sapply function to calculate the sum of all steps in a given day. Below is a histogram of the steps taken by day.


```{r}

completedata <-subset(mydata,complete.cases(mydata)==TRUE) 

bydate<-split(completedata,completedata$date, drop=TRUE)          
totalstepsperday <-sapply(bydate, function(x) sum(x$steps))

mean <- mean(totalstepsperday)
mean <- round(mean)
median <- median(totalstepsperday)

hist(totalstepsperday, main="Steps taken by Day", col="dark green")
```

The mean of the data is `r mean`, and the median of the data is `r median`.

Below is a chart showing the average number of steps by the 5-min interval.  


```{r}
bysteps <- split(completedata,completedata$interval)     
avgofintervalsteps <- sapply(bysteps, function(x) mean(x$steps))    
plot(avgofintervalsteps, type="l",  main="Average of steps by 5-min interval", xlab="Interval", col="cyan", ylab="Average of steps by interval")                         


maxinterval <- names(which.max(avgofintervalsteps))
stepsofmax <- max(avgofintervalsteps)
```


The max steps for 5 minute interval happens at `r maxinterval`.
The average number of steps for that interval is `r stepsofmax`.

To compare the number of steps between weekdays and weekends I created a new data frame.  I then adjusted the date to be a Time class since it originally was read in as a factor class.  Then using weekdays to create a new column that contained the day of the week.  Finally I used a for loop to go through that new day column and change it to either weekend or a weekday.

```{r}
new <- completedata
new$date <- as.Date(new$date, "%Y-%d-%m")
new$day <- weekdays(new$date)
for (i in 1:nrow(new)) {                                      
    if (new[i,]$day %in% c("Saturday","Sunday")) {             
        new[i,]$day<-"weekend"                                 
    }
    else{
        new[i,]$day<-"weekday"                                 
    }
}



```

Below you can see the steps for weekdays which has been plotted in blue.


```{r}

weekday <- subset(new, new$day =="weekday")
weekend <- subset(new, new$day == "weekend")

par(mfrow=c(1,1))  
plot(weekday$steps, type="l",  main="Weekday steps", xlab="Interval", col="blue")
```

It looks different from the steps during the weekends which is plotted below in red. 

```{r}
plot(weekend$steps, type="l",  main="Weekend steps", col="red", xlab="Interval")
```
